### 2016-01-22 03-design.tex
Polish a bit on 02, 04, 04, 05
First half of design edits

The SG-MCMC algorithm described in the previous section has an abundance of
opportunities for parallelism, for the multi-threaded, shared-memory type as
well as for distributed-memory parallelism. The benefit of multi-threaded
parallelism is speed-up of the computation. A distributed implementation
additionally allows to store data in the collective memory of the cluster
machine and increases memory bandwidth which scales with the number of
machines. The downside of a distributed implementation is that it requires
considerably more programming effort.

@@

This section describes how we parallelize the algorithm, which had been
implemented sequentially in~C++.
In most places, the usage of multi-threaded parallelism
is straightforward, therefore, we will discuss details only where appropriate. The
distributed design follows a master-worker paradigm, where in essence
the master controls the parallel operations and the workers perform the
calculations. For thread parallelism, we annotate the program with
OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
MPI~\cite{Forum:1994:MMI:898758}.

@@

The distribution within the graph of the vertices of the mini-batch as well as
the neighbor sets is completely random. That means that there is no locality
in $\pi$ access patterns, especially in the accesses to the neighbor sets
$\Neighbors$.
Hence, there is no opportunity to exploit data locality through caching of
$\pi$.

@@

%\looseness=-1
Modern network technology allows access to the memory of remote machines
through the network cards (NICs) without involvement from the remote host
altogether~\cite{Hamada_infinibandtrade,Beck:2011:PER:2043535.2043537}.
This has opened up the development of
extremely efficient remote memory services, often in the form of DKV stores.
This project stores $\pi$ in a DKV store in the collective memory of the
cluster, where $\pi[i]+\sum\phi[i]$ is the value for key $i$. We decided not
to use any of the existing implementations of RDMA DKV stores, and instead
build our own DKV store on top of the Infiniband ib-verbs~API. The
main consideration is that our use case is unusually simple in a number
of important aspects. The KV layout is static in the sense that there are
no inserts or deletes after the initial population, which allows a static
partitioning of KV pairs over the machines. The values are all of the same size,
a vector of $K+1$ floats. The access pattern is well controlled because the
computation is partitioned into stages that are separated by barriers. A
stage either reads values or updates values. The algorithm ensures that updates
are always targeting unique elements,
therefore, there are no read/write or write/write hazards. The existing DKV
stores (RamCloud~\cite{Ousterhout:2015:RSS:2818727.2806887}, FaRM~\cite{179767},
and reputedly the fastest, Herd~\cite{Kalia:2014:URE:2740070.2626299}) incur
overhead for insert/delete flexibility, concurrency control, or variable-sized
values. Our use case allows us to do any operation in exactly one RDMA read
or RDMA write. Herd argues
that replacing RDMA reads by
an RPC to the server, followed by an unreliable write, is faster. However,
they show this only holds for small payload packets, up to 256B. A $\pi$
packet in our application is typically thousands to hundreds of thousands
of 4-byte floats.

>>>>

The SG-MCMC algorithm described in the previous section has an abundance of
opportunities for parallelism, for the multi-threaded shared-memory type as
well as for distributed-memory parallelism. The benefit of multi-threaded
parallelism is the speedup of the computation. Additionally, a distributed implementation
allows the storage of data in the collective memory of a cluster
and increases aggregate memory bandwidth. Both of these parameters scale with the number of
machines. The downside of a distributed implementation is that it requires
considerably more programming effort.

@@

This section describes how we parallelized the algorithm, which had been
implemented sequentially in~C++.
The usage of multi-threaded parallelism
is straightforward, therefore, we will discuss its details only where appropriate. The
distributed design follows a master-worker paradigm, where
the master controls the sequence of parallel operations and the workers perform the
calculations. For thread parallelism, we annotate the program with
OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
MPI~\cite{Forum:1994:MMI:898758}.

@@

The distribution within the graph of the vertices of the mini-batch as well as
the neighbor sets is completely random. That means that there is \textit{no
locality}
in $\pi$ access patterns, especially in the accesses to the neighbor sets
$\Neighbors$.
Hence, there is no opportunity to exploit data locality through caching of
$\pi$.

@@

%\looseness=-1
Modern network technology allows for Remote Direct Memory Access (RDMA)
through the network cards (NICs) with negligible
overhead~\cite{Hamada_infinibandtrade,Beck:2011:PER:2043535.2043537}. The use
of RDMA enables communicating peers to avoid the high latency associated with
interrupting a processor and waiting for it to act upon a request. Instead, a
process would register a region of memory that can be directly accessed by the
NIC. However, RDMA only provides primitive operations and limited abstractions.
This has opened up the development of
efficient remote memory services, often in the form of distributed key-value
(DKV) stores.

The distributed algorithm implementation stores $\pi$ in a DKV store in the
collective memory of the
cluster, where $\pi[i]+\sum\phi[i]$ is the value for key $i$.
%
We decided not
to use any of the existing implementations of RDMA DKV stores, and instead
build our own DKV store on top of the Infiniband ib-verbs~API.
%
This stems from the fact that our use case is unique in a number
of important aspects. First, the number of keys and values to be stored are
constant throughout the execution of the algorithm. For instance, no records
are dynamically inserted or deleted from the DKV after their initial insertion.
However, the value associated with each key can be updated. Additionally, all
keys are of the same type and represent a unique vertex in the graph.
Therefore, keys can be represented as integer identifiers. Moreover, each key
points to a vector of $K+1$ floats as its value.
%
This data representation can be exploited to simplify the management of
read and write requests. Similarly, the keys can be deterministically
partitioned across the cluster using a fast hash function.

The algorithm's access pattern is well controlled because the
computation is partitioned into stages that are separated by barriers. A
single stage either reads values or updates values. The algorithm ensures that updates
are always targeting unique elements,
therefore, there are no read/write or write/write hazards. The existing DKV
stores (RamCloud~\cite{Ousterhout:2015:RSS:2818727.2806887}, FaRM~\cite{179767},
and Herd~\cite{Kalia:2014:URE:2740070.2626299}) incur
overhead for insert/delete flexibility, concurrency control, or variable-sized
values. Our use case allows us to do any operation in exactly one RDMA read
or RDMA write.
%
Herd argues that replacing RDMA reads by
an RPC to the server, followed by an unreliable write, is faster.
%
However,
the paper shows this only holds for small payload packets, up to 256B. A $\pi$
packet in our application is typically thousands to hundreds of thousands
of 4-byte floats.
